%
% PROJECT: <ETD> Electronic Thesis and Dissertation Initiative
%   TITLE: LaTeX report template for ETDs in LaTeX
%  AUTHOR: Neill Kipp, nkipp@vt.edu
%     URL: http://etd.vt.edu/latex/
% SAVE AS: etd.tex
% REVISED: September 6, 1997 [GMc 8/30/10]
% 

% Instructions: Remove the data from this document and replace it with your own,
% keeping the style and formatting information intact.  More instructions
% appear on the Web site listed above.

\documentclass[12pt]{report}

\usepackage{graphics}
\usepackage{epsfig,dsfont}
\usepackage{times,tikz, tikz-qtree}
\usepackage{amsmath, amssymb, hyperref,float}
%\usepackage[nottoc,numbib]{tocbibind}
\usepackage{framed, pseudocode}
\usepackage{mathrsfs} 
\usepackage{graphicx, dsfont}
\usepackage{amsthm,fullpage}
\usepackage{graphicx,float,hyperref}
\usepackage{verbatim}
\usepackage{multirow} % for column and two spanning in tables
\usepackage{subfigure}
\usepackage[ruled,vlined]{algorithm2e}



\setlength{\textwidth}{6.5in}
\setlength{\textheight}{8.5in}
\setlength{\evensidemargin}{0in}
\setlength{\oddsidemargin}{0in}
\setlength{\topmargin}{0in}

\setlength{\parindent}{0pt}
\setlength{\parskip}{0.1in}
%%%%----------------------------------------- Code for notation list
% Macro for 'List of Symbols', 'List of Notations' etc...
\def\listofsymbols{\input{symbols} \clearpage}
\def\addsymbol #1: #2#3{$#1$ \> \parbox{4.5in}{#2 \dotfill \pageref{#3}}\\} 
\def\newnot#1{\label{#1}} 
%%%%%--------------------------------------End of code for notation list


%%%%----------------------------------------- Code for abbreviation list
% Macro for 'abbreviations' etc...
\def\listofabbreviations{\input{abbreviations} \clearpage}
\def\addabbrev #1: #2#3{#1 \> \parbox{6in}{#2 \dotfill \pageref{#3}}\\} \def\newabbrev#1{\label{#1}} 
%%%%%--------------------------------------End of code for notation list


\newtheorem{mydef}{Definition}
\newtheorem{lem}{Lemma}
\newtheorem{thrm}{Theorem}

\graphicspath{ {/Users/rlucas7/Dropbox/Public/Thesis_Proposal/figures/} }
%\graphicspath{ {images_folder/}{other_folder/}{third_folder/} } use this if there are several filepaths or folders for images 

%for fancy script letters... $\mathscr{ABCDEFGHIJKLMNOPQRSTUVWXYZ}$
\newcommand{\Risk}{(1/n)\sum_{i=1}^n\mathbb{I}(g(Z_i)=y_i)}
\newcommand{\erisk}{(1/n)\sum_{i=1}^n\mathbb{I}(\hat{g}(Z_i)=y_i)}
\newcommand{\bayesorprob}{Bayesian}
\newcommand{\yvec}{\underline{y}}
\newcommand{\xvec}[0]{\underline{x}}
\newcommand{\simp}[1]{\mathbb{S}^{#1}}
\newcommand{\Rsp}[1]{\mathbb{R}^{#1}}
\newcommand{\muvec}[0]{\underline{\mu}}
\renewcommand{\vec}[1]{\underline{{#1}}}
\newcommand{\iid}[0]{$i.i.d.$}
\newcommand{\invgam}[3]{\text{Inv-Gamma}({#1}\vert{#2},{#3})}
\newcommand{\Sbeta}[4]{S\beta(#1\vert#2, #3, #4)}
\newcommand{\ALT}{ALT}
\newcommand{\ALN}{ALN}
%\usepackage[mathscr]{euscript} 
% for less fancy script letters... $\mathscr{ABCDEFGHIJKLMNOPQRSTUVWXYZ}$

% The proposal format is the following:
%        30 pages
%        12 point type
%        1 inch margins all around = 6.5   inch columnT
%        (Total:  30 * 6.5   = 195 page-inches)
%
% For letter-size paper: 8.5 in x 11 in
% Latex Origin is 1''/1'', so measurements are relative to this.

\linespread{1.6} % 1.6 = double space, 1.3 = 1.5 space comment out for regular


\topmargin      0.0in
\headheight     0.0in
\headsep        0.0in
\oddsidemargin  0.0in
\evensidemargin 0.0in
\textheight     9.0in
\textwidth      6.5in


%\title{{\bf Doctoral Thesis } \\
%\it Submitted to the faculty at Virginia Tech}
%\author{ {\bf Lucas Roberts}  \\
%Department of Statistics \\
%Virginia Tech \\
%{\small rlucas7@vt.edu}
%}
%\date{\today}




\begin{document}

\thispagestyle{empty}
\pagenumbering{roman}
\begin{center}

% TITLE
{\Large 
Variable Selection and Decision Trees: The DiVaS and ALoVaS Methods
}

\vfill

Lucas Roberts

\vfill

Dissertation submitted to the Faculty of the \\
Virginia Polytechnic Institute and State University \\
in partial fulfillment of the requirements for the degree of

\vfill

Doctor of Philosophy \\
in \\
Statistics

\vfill

Scotland C. Leman (Chair) \\
Leanna L. House \\
Chris L. North \\
Eric P. Smith

\vfill

September 24, 2014 \\
Blacksburg, Virginia

\vfill

Keywords: Statistics, Decision Trees, Variable selection, Additive Logistic Normal
\\
Copyright 2014 Lucas Roberts

\end{center}

\pagebreak

\thispagestyle{empty}
\begin{center}

{\large Variable Selection and Decision Trees: The DiVaS and ALoVaS Methods
}

\vfill

Lucas Roberts

\vfill

(ABSTRACT)

\vfill

\end{center}



%\pagebreak
%\begin{abstract}
In this thesis we propose a novel modification to Bayesian decision tree methods. We provide a historical survey of the statistics and computer science research in decision trees. Our approach facilitates covariate selection explicitly in the model, something not present in previous research. We define a transformation that allows us to use priors from linear models to facilitate covariate selection in decision trees. Using this transform, we modify many common approaches to variable selection in the linear model and bring these methods to bear on the problem of explicit covariate selection in decision tree models. We also provide theoretical guidelines, including a theorem, which gives necessary and sufficient conditions for consistency of decision trees in infinite dimensional spaces. Our examples and case studies use both simulated and real data cases with moderate to large numbers of covariates. The examples support the claim that our approach is to be preferred in large dimensional datasets. Moreover, our approach shown here has, as a special case, the model known as Bayesian CART.    

%The thesis proposal is a type of contract between the faculty and the student. 
%An accepted thesis proposal indicates that the work proposed by the student, 
%once completed, will be accepted by the faculty as sufficiently innovative and 
%substantial as to be recognized with the award of the degree. It is part of 
%the training of the student's research apprenticeship that the form of this 
%proposal must be as concise as those proposals required by major funding 
%agencies.

%\end{abstract}

\pagebreak 

{\large \textbf{Acknowledgements}
}

The road to completing this thesis was not easy at times and I am grateful for the people who helped me along the way. 

I would like to thank Dr. Scotland Leman for entertaining my ideas and providing encouragement throughout all aspects of my research. I would like to thank Dr. Leanna House for teaching me the latent variable trick in her hierarchical modeling class. I'd like to thank Dr. Eric Smith for showing me what it means to think like a scientist and how to ask really good applied questions. Thank you Dr. Chris North for providing an alternate viewpoint on many aspects of the research. Also, thank you Dr. Birch and the other people involved in the decision to accept me into the PhD program. 

Thank you to Dr. Gill Giese for teaching me about grapes, wine, and applied research. Peter Boyer and Martin Ahrens at Capital One were responsible for introducing me to decision tree models in the first place. Jim King, at The Hartford, showed me how Amazon web services work and how EC2 can provide cost effective computational solutions to problems. 

Thank you to Janice and Jim Burrows, Adam and Martha Sanders, Courtney and John Scherer, Irina and Bogdan Metehoiu, and Perry Gibson for all their support along the way. 

Finally, I would like to thank Dr. Denisa Anca Olteanu Roberts, my partner in all things, without whom none of this would have been possible. Through good times and bad she held the line and encouraged me to continue, she proofread and asked insightful questions, and she showed me other cultures and continents. She showed me what it means to think like an economist, a computer scientist, and a statistician, and that learning does not stop even after completing a PhD. For all this and more, thank you.  

\pagebreak
\tableofcontents
\pagebreak

\listoffigures
 \addcontentsline{toc}{chapter}{\textbf{List of Figures}}
\pagebreak

\listoftables
 \addcontentsline{toc}{chapter}{\textbf{List of Tables}}
\pagebreak

\clearpage
\phantomsection
 \addcontentsline{toc}{chapter}{\textbf{List of Symbols}}
\listofsymbols
\pagebreak

\clearpage
%\phantomsection
 \addcontentsline{toc}{chapter}{\textbf{List of Abbreviations}}
\listofabbreviations
\pagebreak

\cleardoublepage
\pagenumbering{arabic}

% intro chapter
\input{intro_chapter.tex}

% Chapter on Decision trees and on Bayes Decision trees
\input{overview_chapter.tex}

%% Chapter on DiVaS method 
\input{dirichlet_variable_selection.tex}

%% Chapter(s) on ALoVaS method

\input{ALOVAS.tex}

\newpage
\input{alovas_ex_and_case_study}

\newpage
\input{synthesis.tex}
\newpage
\input{discussion.tex}
\pagebreak

\input{appendixA.tex}

\pagebreak

%\cleardoublepage

%\addcontentsline{toc}{chapter}{Bibliography}
%\bibliographystyle{abbrv}
%\addtocontents{toc}{\protect\setcounter{tocdepth}{-1}}% - See more at: http://www.howtotex.com/general/control-your-table-of-contents/#sthash.28ZEx8jt.dpuf
%\phantomsection
\bibliographystyle{abbrv}
\bibliography{tree_biblio,tree_biblio1,tree_biblio2}


\end{document}


